A simulation study of gene flow and local polygenic adaptation motivated by threespine stickleback evolution

### 2018-10-22 Stickleback_Paper.tex
Update on Overleaf.

\title{The sweet spot for local polygenic adaptation: a case study motivated by stickleback}
\author{Jared Galloway, William A. Cresko, and Peter Ralph}
\maketitle

@@

How do stickleback share adaptations between lakes? A simulation study.

A simulation study of gene flow and local polygenic adaptation

A simulation study of gene flow and local polygenic adaptation in stickleback fish

effect loci (locus) | QTL | causal: = ? REGIONS UNDERLYING TRAIT VALUE , GENES, ---

\section*{Abstract}

Threespine stickleback fish provide one of the most striking examples of local adaptation. This hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations
that have apparently repeatedly adapted to freshwater conditions
often using the same genetic basis.
In this paper, we use simulations motivated by stickleback populations
to examine what amounts of gene flow favor stable metapopulation polymorphism
with allele sharing,
and to further dissect the underlying dynamics.
We find that rapid, repeated adaptation 
using alleles maintained at low frequency by migration-selection balance
(the ``transporter hypothesis'')
occurs over a realistic range of intermediate rates of gene flow,
between slow, independent adaptation with low gene flow
and large migration load at high gene flow.
Rapid adaptation occurs when the freshwater haplotype has been maintained,
through historical introgression, in marine environments without tight linkage to 
marine adapted alleles.
%This is mediated mainly by the total downstream influx of alleles \emph{(maybe??)(No)}.
%We do not see evidence for strong effects encouraging genomic clustering of causal alleles
%or towards particular dominance coefficients.
$F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow,
The results support existing theory of local adaptation,
and provide a more concrete look at a particular, empirically motivated example.

\section*{Introduction}

@@

\emph{intro outline:}
Start off with stickleback: repeated, rapid, parallel adaptation with obvious phenotypes.
But, most species don't seem to have such obviously structured polymorphism: 
what's special about stickleback?
The transporter hypothesis is a mechanism; what's required to make it work?
Other species adapt rapidly in similar ways:
e.g., Darwin's finches and island mice;
do these differ through not having such a big, connected reservoir of genetic variation?
or by not having such a large total population size of the ``other'' habitat?
Or, is this situation more common than we know, with less obvious phenotypes?
Also, can we actually tell what's happening from genetic data?

Recently, multiple instances of similar (parallel) underlying genetic basis of rapid local adaptation has brought about 
questions surrounding the origins and maintenance of genetic polymorphism.
A empirical model is the Alaskan populations of freshwater and marine Ninespine Stickleback fish.
In 1964, The Great Alaskan Earthquake caused an uplift of Middleton island and in turn,
introduced a group of freshwater ponds around the perimeter of the island. 
Quickly inhabited by the surrounding marine population of Stickleback,
\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be 
parallel to freshwater stickleback that have been separated for over $13,000$ years.
In freshwater stickleback, the number of lateral plates are reduced
and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.
% \jgg{Bill - more specifics about Bill, Suzie, Kristin  \& or Thom's papers .. ? }
These results leave us with questions surrounding the nature of rapid adaptation.
Does convergent evolution breed it's own solution (haplotype) for every new selective pressure, 
or can these solutions can be efficiently shared across multiple sub - populations facing similar selective pressures.

The leading hypothesis for stickleback is that rather than acting on new mutations, 
adaptation to freshwater environments is sped up through selection on standing genetic variation (SGV) found in marine populations. 
\jgg{Bill: maybe you could make this section a little more specific}
One clear example of this is the gene \textit{eda} 
which has been shown to regulate the number of lateral plates. 
While this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently.
Novel evidence from natural populations has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.
\citet{schluter2009genetics} 
suggested the ``transporter''-hypothesis.  
This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
The continued selection on freshwater favored alleles and introgression between the two sub-species,
allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection reconstructs the freshwater haplotype \citep{schluter2009genetics}.
An alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, 
and their haplotype is passed down directly. 
This hypothesis has been shown to be unlikely due to finding a high frequency of freshwater alleles in the ocean,
and almost no freshwater individuals.

\plr{Bring up the possibility of more than one freshwater haplotype here.}
\jgg{Re: should I? Doesn't Thom's work and parallel above suggest they have the same alleles?}

If selection on standing genetic variants is key in rapid adaptation, 
many questions are exposed concerning the surrounding population genetics parameters and underlying genomic architecture.
%How rapidly can selection act on standing genetic variants at a given value of migration, $M$?
What scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation
of a population derived from $X$ to an introduced environment with selective pressure $Y$?
What is the origin of pre-existing adaptive variants?
How are the variants structured, historically and across the genome underlying the trait?
%How does population genetics data influenced by historical introgression and selection on standing genetic variants?
%How many alleles underly a certain trait and how many are swept away by drift? 
Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data.
Many biologists today make use of genome wide association studies (GWAS) and 
$F_{st}$ across the genome to estimate regions responsible for certain traits. 
Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples. 
This being the case, what are the biases we can expect to see in real data for certain levels of introgression.

There are many different parameters that carry a large impact on the questions above.
\citet{ralph2015convergent} 
explored the dichotomy of selection on new alleles and those brought in by migrants from a similar selective pressure.
Answering the fundamental questions surrounding spatial resolution of convergent evolution. 
%This lead to an understanding of the process by which alleles move from an existing 'patch' to an introduced 'patch' with the 
%same selective pressure. 
\jgg{It seems like this could be a good parallel to what we're doing here, or maybe for discussion?}

In this paper, we explore time to local adaptation as a function of migration rate through an environment (marine) with opposing selective pressure. 
Next, we compare to real data and explore how $F_{st}$ data can be impacted by introgression.
%Explain what we do and what we find that is important

%We use forward moving simulations to model marine and freshwater populations of stickleback.
%We then observe the effect (focusing on migration and recombination, for now) 
%of different parameters on local adaptation of introduced freshwater environments.

%%%%%%%%%%%%%%%%
\section*{Methods}

We explored these questions
using forwards-time simulations with explicit genomic representation of a quantitative trait
in SLiM \citep{haller2017slim}.
The details (described below) are motivated by previous population genomic studies of threespine stickleback,
but remain simplistic in some aspects due to computational constraints.
\jgg{should we brag about treeSeq here a little?}
Possibly the most important caveat is that simulated population sizes are much smaller
than the census size of the Threespine Stickleback population;
we will consider the likely effects of this in the Discussion.
\plr{TODO}

\paragraph{Habitat and geography}
Our model has two habitat types, which are considered separate populations
with a fixed total size of 5,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes.
The scheme, depicted in Figure \ref{fig:Geo},
roughly models a set of freshwater habitats
along a stretch of coastline.
The ``marine'' habitat  is a continuous, one-dimensional range of 25 units of length,
and the ``freshwater'' habitat is divided into ten subpopulations
(which we call ``lakes''), each connected to the marine habitat
at positions $i - 1/2$ for $1 \le i \le 25$.

Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in the two different habitats. This situation roughly models the cumulative effect of the various phenotypes
such as armor morphology, body size, craniofacial variation and opercle shape (among others)
on which divergent selection acts in the two environments.
The optimal trait value in the marine habitat is $+10$,
and in the freshwater habitat is $-10$. Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is
Gaussian with standard deviation, i.e.,
\begin{align*}
    f(x; x_0)
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
  		\caption{
			A representation of the geographic and evolutionary history of all populations throughout the simulation. 
		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 25.0]. 
		Each freshwater lake$_{i}$ for both the introduced and freshwater populations
		is a discrete population connected by migration only through the marine, at position $i - 0.5$. 
		Each of the introduced lakes have the same spatial dynamic and selective pressure as the pre-existing lakes, but arise at 100K generations into the simulation.
		The marine has $5000$ diploid individuals while each of the lakes fluctuate  $200$ diploid individuals. 
		the introduced population is initialized as a copy of all marine individuals to model marine 
		individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
		We then observe the selection process of marine individuals and following generations 
		with the new selective pressure across a range of parameter values. 
			}
  		\label{fig:Geo}
	\end{center}
\end{figure}

\paragraph{Potential genomic architecture}\bill{Initial genomic architecture? The header is a little confusing to me}
Each individual carries two copies of a linear chromosome of $10^8$ loci representing a realistic size genome.
New mutations occur with probability $10^{-10}$ per locus per generation in regions affecting phenotype and $10^{-8}$ in neutral regions.
Ten ``effect regions'' of $100 Kb$  loci each are spread evenly along the chromosome; 
mutations in these regions have the potential to affect trait value of the individual which posses them
Each mutation in these regions is either additive, completely recessive,
or completely dominant (with equal probability).
Effect sizes are chosen randomly from an exponential distribution with mean $1/2$,
either positive or negative with equal probability.
Between the effect regions is $9.9 Mb$ of neutral loci.
Individual trait values are determined additively from the diploid genotypes.
Concretely, an individual that is heterozygous and homozygous for mutations
at loci $H$ and $D$ respectively has trait value
$x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$,
where $s_\ell$ is the effect size of the mutation at locus $\ell$.
Subsequent mutations at the same locus replace the previous allele.

\paragraph{Life cycle}
Each generation, the two parents of each new offspring are chosen 
proportional to their fitness (all individuals are hermaphroditic),
and the contributing genomes are produced by Poisson recombination
with an average of one crossover per chromosome per generation
($10^{-8}$ per locus per generation).
The model is Wright-Fisher, so that each generation, in each habitat,
5000 new individuals are produced,
and so to keep population sizes roughly constant within each lake,
at the start of each generation we rescale fitness values in the freshwater habitat
so that the sum of the fitness values within each lake are equal.

Migration occurs both locally along the coastline in the marine habitat
and between the marine habitat and the lakes,
and can be thought of as occurring at the juvenile stage.
The between-habitat migration rate is denoted $m$,
and will be called simply the ``gene flow''.
$\approx m$ individuals \emph{per} lake \emph{per} generation, have parents from the opposing selective pressure.
A first parent is chosen proportional to fitness,
and a mate is chosen from the same lake as the first, also proportional to fitness.
The resulting offspring is given a spatial location in the marine habitat
at the location of the parent's lake.
Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$):
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, 
also proportionally to fitness but reweighed by a Gaussian function 
of the distance separating the two, with standard deviation $1/2$. 
More specifically, if the first parent is marine individual $i$,
then marine individual $j$ is chosen as the mate
with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations.
Finally, each new marine offspring is given a position 
displaced from the first parent's position by a random Gaussian distance
with mean 0 and standard deviation 0.02, and reflected to stay within the population.
New offspring in the freshwater habitat are chosen in the same way,
except the probability that the parents are marine individuals is $m$;
any new freshwater offspring produced by marine individuals
are assigned to the lake nearest to the position of the first marine parent.

\paragraph{New lakes}
To study adaptation in newly appearing freshwater habitats,
we introduce a new set of lakes midway through the simulation.
As with the old lakes, these are a set of discrete demes connected by migration \emph{only} to the marine.
The initial generation of individuals in these new lakes were created as a copy of the marine individuals 
at the generation of introduction.
The new lake's geography acts like an independent copy of the original set of lakes -- in particular,
the two sets of lakes each have 5,000 individuals at all times 
(Gene flow constricted to within the lakes and through migration to the marine environment).
Since this doubles the number of lake-to-marine immigrants,
after this happens the probability that a new marine individual has freshwater parents
is $2m$ instead of $m$.

\paragraph{TreeSeq}
This case study involves analyzing large simulations which model populations under the influence of selection. 
To fully understand the origins of freshwater alleles, calculate statistics, and avoid the computationally expensive task of 
simulating neutral mutations, we have used SLiM's \emph{treeSeq} feature to track the genealogical history of all samples throughout the simulation. 
This feature outputs a compact representation the genomic outcome of every meiosis relevant to tracked samples (Often thought of as the ARG).

Using the output tree sequence, 
we were able to classify each freshwater allele in each genome of the new lakes population, at time of adaptation,
as deriving from one of four origins:
(1) Migrants, brought in from a migrant after the introduction of the new lakes 
(2) Pre-existing, passed down from an individual in the original generation of the new lakes, 
however, not defined to be an adaptive variant in the original lakes
(3) Captured, Pre-existing, passed down from an individual in the original generation of the new lakes, 
and defined to be adaptive in the original lakes
(4) De novo, mutated and propagated from an individual existing in the introduced lakes population, after introduction. 
\jgg{We should discuss further what these categories aught to be named}

\begin{comment}
We also used the tree sequence to count the "fitness" of individuals in the initial generation of the new lakes. 
The "fitness" of an individual in this case, is defined to be the total number off eventual offspring, in the generation at the time of adaptation, 
which have inherited a adaptive variant from that individual. 
To calculate this, we only considered gene trees which overlapped an adaptive variant site (defined at time of adaptation).
We then counted the total number of offspring which had inherited from any adaptive variant site, 
for each individual in the initial generation. 
\end{comment}

To determine the ability for the new lakes to select upon the standing variation in the marine at different rates of historical introgression,
We were able compute the trait value of all effect regions in the initial population.
Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
we determined the expected total effect sizes of the haplotypes that fix.
Concretely, it is the sum across the 10 effect regions of
$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.

\paragraph{Descriptive statistics}\bill{Population genetic analyses? Instead of the more general 'Descriptive statistics'?}
To assess whether new lakes adapt using existing genetic diversity,
we define a \emph{freshwater allele}
to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
while remaining lower than $0.5$ in the marine. 
This categorization is made for each generation using the allele frequencies from that generation,
and so changes with time.
Alleles common in the newly introduced lakes do not count
if they are not also common in the original lakes.
They are defined this way because the transportation hypothesis 
does not specify where or when an advantageous mutation arises,
but simply suggests that any sufficiently common freshwater adapted allele 
could participate in adaptation in new habitat \citep{schluter2009genetics}.

\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, 
is defined to be the generation at which 
the difference between the average trait value in the original and the introduced freshwater populations 
is less than 0.5.

We describe overall genetic differentiation between the habitats using $F_{ST}$,
calculated on a per-locus basis.
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
in the freshwater and marine habitats, respectively,
and $\bar p = (p_f + p_m)/2$,
then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.

\section*{Results}

We varied the migration rate, $m$ across separate simulations
from $5 \times 10^{-5}$ to $5 \times 10^{-1}$ by steps one order of magnitude.
From here on out we will refer to these rates of gene flow as running simulations ranging from {0.01 - 100.0}
migrants \emph{per} lake \emph{per} generation.
%i.e., between 0.01 and 100 migrants per generation to and from each of the freshwater lakes.
This had a strong effect on many aspects of adaptation,
including how fast adaptation occurred in each lake,
how many alleles were shared between lakes,
and the population genetic signals left behind.
We find a threshold of gene flow which allows 
the freshwater haplotype to be maintained in the marine without creating linkage to marine alleles. 
below this threshold, selection on pre-existing freshwater alleles to be much less efficient 
due to background selection. \jgg{did I get this right?}

Interestingly, all rates of $m$ aside from the two highest had little affect on populations ability to locally adapt.
All rates of gene flow aside from the lowest, helped both the old lake's and the new lake's freshwater populations share pre-existing freshwater adapted alleles.
The sharing of alleles at $m = 1.0$ resulted in rapid adaptation ($\approx 32$ generations) of the introduced population to adapt to the freshwater environment.
At the lowest migration rate, lakes adapted nearly independently,
while at the highest migration rate, the habitats act like a single metapopulation
with substantial migration load.
We will now dissect what happens between these extremes.

@@

\begin{figure}
	\begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
  		\caption{ 
        		Mean individual trait values in the marine habitat (blue line),
        		the original lakes (light green lines; average in dark green),
        		and the introduced lakes (yellow lines; average in orange)
        		across the course of two simulations, with migration rates of
        		\textbf{(top)} $m=5 \times 10^{-5}$ and
        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation individual
        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
        		Optimal trait values in the two habitats are at $\pm 10$.
		}
  		\label{fig:phenotype_ts2}
	\end{center}
\end{figure}

Local adaptation occurred in all simulations:
as shown in Figure \ref{fig:phenotype_ts2},
freshwater and marine populations diverged
until the trait means were close to the optimal values in each habitat. 
The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
as jumps in the mean trait value;
in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations,
\jgg{I'm not sure what this means? how do we know how long the jumps take?}
and move the trait by an amount of order 1.
Trait variation within each population was small compared to the difference between populations,
with interquartile ranges of around XXX.\plr{TODO}
Across all parameter values, differences at around 16 commonly polymorphic sites 
(eight the shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.

>>>>

\title{A simulation study of gene flow and local polygenic adaptation motivated by threespine stickleback evolution}
\author{Jared Galloway, William A. Cresko, and Peter Ralph}
\maketitle

@@

\textbf{JARGON WE SHOULD THINK ON ??}

effect loci (locus) | QTL | causal: = ?

Original Lakes | initial | $P_{1}$, $P_{3}$, $P_{3}$ = ?

Migration Rate | Historical Introgression = ?

pre-existing freshwater adapted alleles | standing genetic variants = ?

\textbf{Pending things to do?} that maybe we do, maybe not

longer genomes ($10^8$ not $10^5$ but rescaling everything accordingly)

verify that adaptation of introduced pops is due to initial diversity not subsequent migration

\section{Abstract}

Threespine stickleback fish provide a striking example of local adaptation in the context of recurrent gene flow. This Northern hemisphere-wide metapopulation includes both marine populations and a large number of smaller freshwater populations that have repeatedly adapted to freshwater conditions, often by using standing genetic variation. Here we use genealogical simulations to determine the levels of gene flow can best match the observed patterns of allele sharing among habitats in stickleback, and to provide a framework for better understanding of the dynamics of gene flow and local adaptation for the maintenance and reuse of standing genetic variation. We find that rapid, repeated adaptation using alleles maintained at low frequency by migration-selection balance occurs over a realistic range of intermediate rates of gene flow.  Slow and independent adaptation occurs with low gene flow, whereas large migration load occurs at high gene flow, mediated mainly by the total downstream influx of alleles \emph{(maybe??)}. We do not see evidence for strong effects encouraging genomic clustering of causal alleles or towards particular dominance coefficients. $F_{ST}$ scans for adaptive alleles are more likely to succeed with higher rates of gene flow, The results support existing theory of local adaptation, and provide a more concrete look at a particular, empirically motivated example.

\textbf{**We need to be more crystal clear in the abstract about the findings from the paper**}

\section{Introduction}

@@

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. While it has proven valid in numerous studies in the field and the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. A growing number of studies have identified instances of convergent or parallel adaptive evolution that utilize standing genetic variation. What evolutionary processes promote the maintenance of re-use of such genetic polymorphisms? How do these processes interact with diverse geographic contexts, as well as variation in genetic processes such as recombination, to change the likelihood of so-called standing genetic variation (SGV).

Clues come from population genomic studies in the wild that have recently been powered by advances in sequencing technologies. One such organism is the threespine stickleback. This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution. The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere, which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences. This adaptive phenotypic evolution to freshwater can occur in mere millenia, in many cases since the end of the last ice age about 13,00 years ago. Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, often involving very similar phenotypic shifts as seen in contemporary populations.

A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution. These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow. This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades. The answer is a clear yes. For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. Quickly inhabited by the surrounding marine population of stickleback, \citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years. In freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows the same expansion of the dorsal region and reduction of the ventral region.

 But how can evolution occur so quickly? The leading hypothesis for stickleback is that rather than acting on new mutations, adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates. While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently. More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that most regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}. Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.

These empirical data generally support the ``transporter''-hypothesis proposed by Conte and Schluter in 2009 \citet{schluter2009genetics}, which is a conceptual model for the flow of freshwater alleles from multiple smaller freshwater populations into much larger and less structured marine populations through hybridization events. Alleles in this asymetrically structured metapopulation can then be recycled for subsequent freshwater adaptation. Several questions remain, however, about the manner and degree to which these alleles are scattered among the marine fish and must be re-assembled. Is it more akin to the atomization and rebuilding that occurred in the Star Trek series that motivated this hypothesis? Alternatively does it occur in more of a patchwork by individuals, or their early generation hybrids, in geographically adjacent freshwater habitats moving quickly through the marine environment to seed new freshwater populations. More generally, how do different levels of gene flow and local adaptation interact with the asymmetric nature of stickleback population structure affect the dynamics of the origin and rate of adaptive alleles, as well as their eventual organization into marine and freshwater typical genomes?

Here we develop a forward simulation approach in SLiM to model the stickleback population structure in marine and freshwater habitats to document the effect of variation in gene flow and recombination on the genetic and genomic architecture of local adaptation. We address how rapidly can selection act on standing genetic variants at a given value of migration, $M$, and what the precise origin is of those variants. We ask what scale of historical introgression between populations with selective pressures $X$ and $Y$, allows for rapid and parallel local adaptation of a population derived from $X$ to an introduced environment with selective pressure $Y$? How are the variants structured, historically and across the genome underlying the trait? Furthermore, how can we infer causal loci for regions of the genome that must be driving the rapid adaptation, from real data. Many biologists today make use of genome wide association studies (GWAS) and $F_{st}$ across the genome to estimate regions responsible for certain traits. Unfortunately, the data can be heavily influenced by the level of introgression and population structure of the samples. This being the case, what are the biases we can expect to see in real data for certain levels of introgression.

We find that .... \textbf{1. rewrite the previous paragraph to be more concise and tight, and 2. briefly restate our major findings which should line up well with those presented in the abstract. The following I grabbed from the beginning of the Results section}
*****
We varied the migration rate, $m$, from $5 \times 10^{-5}$ to $5 \times 10^{-2}$ per individual, i.e., between 0.01 and 10 migrants per year to and from each of the ten lakes. This had a strong effect on many aspects of adaptation, including how fast adaptation occurred in each lake, how much alleles were shared between lakes,
and the population genetic signals left behind. At the lowest migration rate, lakes adapted nearly independently, while at the highest migration rate, the habitats are beginning to act like a single metapopulation with substantial migration load. We will now dissect what happens between these extremes. Interestingly, the ability for separate populations to locally adapt to their own selective pressure was relatively unaffected until the highest rate of migration between the marine and freshwater environments. All rates of historical introgression aside from the lowest, helped both the initial and introduced freshwater populations share pre-existing freshwater adapted alleles. The sharing of alleles resulted in rapid adaptation ($\approx 100$ generations) of the introduced population (split from the marine population) to adapt to the freshwater environment. We also found that larger amounts of migration allow for more statistical power and less false positives in the resulting population genetics data ($F_{st}$ per SNP across the genome).
****

%%%%%%%%%%%%%%%%
\section{Methods}

We explored these questions using forwards-time simulations with explicit genomic representation of a quantitative trait in SLiM \citep{haller2017slim}. The details (described below) are motivated by previous population genomic studies of threespine stickleback, but remain simplistic in some aspects due to computational constraints.
Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback population;
we will consider the likely effects of this in the Discussion. \plr{TODO}

\paragraph{Habitat and geography}
Our model has two habitat types, which are considered separate populations with a fixed total size of 2,000 diploid individuals each. The marine population is treated as a single large deme, but the freshwater population is subdivided into local subpopulations or demes. The scheme, depicted in Figure \ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. The ``marine'' habitat  is a continuous, one-dimensional range of 10 units of length, and the ``freshwater'' habitat is divided into ten subpopulations (which we call ``lakes''), each connected to the marine habitat at positions $i - 1/2$ for $1 \le i \le 10$.

Fitness in each habitat is determined by a single continuous trait with an inherent tradeoff such that alternative values of the trait have high or low fitness in tht two different habitats. This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, body size, craniofacial variation and opercle shape (among others) on which divergent selection acts in the two environments. The optimal trait value in the marine habitat is $+10$, and in the freshwater habitat is $-10$. Fitness of a fish with trait value $x$ in a habitat with optimal value $x_\text{opt}$ is Gaussian with standard deviation, i.e.,

\begin{align*}
    f(x; x_0)
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{GeographyDiagram}
  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. The marine is a continuous population with spatial positions ranging from [0.0 - 10.0]. Each freshwater lake$_{i}$ for both the introduced and freshwater populations is a discrete population connected by migration only through the marine, at position $i - 0.5$. The introduced lakes have the same spatial location (but separate?) and selective pressure as the original lakes, but arise at 40K generations. The marine has $2000$ diploid individuals while each of the lakes fluctuate around $200$. The introduced population is initialized as a copy of all marine individuals to model marine individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island. We then observe the selection process of marine individuals and following generations with the new selective pressure across a range of parameter values. 
		%Migration between the marine and original lakes is 
		%set to $0$ the generation before introduction to avoid direct migrants from original $->$ introduced lakes.
		}
  		\label{fig:Geo}
	\end{center}
\end{figure}

\paragraph{Initial genomic architecture} Each individual carries two copies of a linear chromosome of $10^5$ loci (which could be thought of roughly as a kilobase each). New mutations occur with probability $10^{-7}$ per locus per generation. Ten ``effect regions'' of 100 loci each are spread evenly along the chromosome (separated by 4950 loci) in which mutations affect the trait; mutations in other parts of the genome are neutral. Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability), and with an effect sizes chosen randomly from an Exponential distribution with mean $1/2$, either positive or negative with equal probability. Individual trait value are determined additively from the diploid genotypes. Concretely, an individual that is heterozygous and homozygous for mutations at loci $H$ and $D$ respectively has trait value $x = \sum_{\ell \in H} h_\ell s_\ell + \sum_{\ell \in D} s_\ell$, where $s_\ell$ is the effect size of the mutation at locus $\ell$. Subsequent mutations at the same locus replace the previous allele.

\paragraph{Equilibrium in initial marine and freshwater populations} Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic),and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation). The model is Wright-Fisher, so that each generation, in each habitat, 2000 new individuals are produced, and so to keep population sizes roughly constant within each lake, at the start of each generation we rescale fitness values in the freshwater habitat so that the sum of the fitness values within each lake are equal.

Migration occurs both locally along the coastline in the marine habitat and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. The between-habitat migration rate is denoted $m$, and will be called simply the ``migration rate''. Each new individual in the marine habitat has freshwater parents with probability $m$; to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen,  also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with stanadard deviation $1/2$.

Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_i) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations. Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, and reflected to stay within the population. New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; any new freshwater offpsring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.

\paragraph{Population genomic analyses of new freshwater populations} To study adaptation in newly appearing freshwater habitats, we introduce a new set of lakes midway through the simulation. The initial set of individuals have parents chosen as above from the marine habitat, and act like an independent copy of the original set of lakes -- in particular, the two sets of lakes each have 2,000 individuals at all times. Since this doubles the number of lake-to-marine immigrants, after this happens the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.

To assess whether new lakes adapt using existing genetic diversity, we define a \emph{pre-existing freshwater adapted allele}
to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes, while remaining lower than $0.5$ in the marine. This categorization is made for each generation using the allele frequencies from that generation, and so changes with time. Alleles common in the newly introduced lakes do not count
if they are not also common in the original lakes. They are defined this way because the transportation hypothesis does not specify where or when an advantageous mutation arises, but simply suggests that any sufficiently common freshwater adapted allele could participate in adaptaiton in new habitat \citep{schluter2009genetics}.

\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, is defined to be the generation at which the difference between the average trait value in the original and the introduced freshwater populations is less than 0.5. We describe overall genetic differentiation between the habitats using $F_{ST}$,
calculated on a per-locus basis. Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively,
and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.

%%%%%%%%%%%%%%%%%
\section{Results}

@@

% The progression of adaptation. How you can see introduction new alleles

Local adaptation occurred in all simulations: as shown in Figure \ref{fig:phenotype_ts2}, freshwater and marine populations diverged until the trait means were close to the optimal values in each habitat. (Remaining simulations are shown in Figures \ref{fig:phenotype_ts1} and \ref{fig:phenotype_ts4}.) Greater migration decreased adaptedness only weakly except for at the highest migration rate (10 migrants per generation per lake), where the mean trait value equilibrated at roughly half its optimal value (Figures \ref{fig:MeanPhenotype} and \ref{fig:phenotype_ts4}). The establishment of new alleles in the lakes is visible in Figure \ref{fig:phenotype_ts2}
as jumps in the mean trait value; in the two simulations these occur on a time scale of 100 (lower migration) to 1000 (higher migration) generations, and move the trait by an amount of order 1. Trait variation within each population was small compared to the difference between populations, with interquartile ranges of around XXX.\plr{TODO} Across all parameter values, differences at around 16 commonly polymorphic sites (eight the shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats.

\plr{ALSO TODO: CHECK THIS}
